Index measures for oak decline severity using phenotypic descriptors

### 2019-12-14 manuscript/manuscript.Rmd
update PDI response surfaces method. Re-arrange of methods and results sections

### Descriptor contribution decline indexes

@@

For the PDI, missing crown and crown transparency were compared for example cases of a healthy, a moderate and a severly tree.
The example cases were identified by selecting the descriptor values for trees with PDI values closes to the 25, 50 and 75% quantiles for the healthy, moderate and severely declining trees respectively.
The value ranges for the missing crown (`r glue("{min(site_corrected_analysis_suitable_data[['Missing crown (%)']])}-{max(site_corrected_analysis_suitable_data[['Missing crown (%)']])}%")`) and crown transparency (`r glue("{min(site_corrected_analysis_suitable_data[['Crown transparency (%)']])}-{max(site_corrected_analysis_suitable_data[['Crown transparency (%)']])}%")`) descriptors divided into 100 increments.
For a given case, each combination of theses simulated descriptor values were used to calculate a PDI value from the supervised random forest regression model, building a model response surface.

@@

### Overall descriptor contributions and trends to the decline indexes

- Top ranked descriptors contributing to the decline indexes
- Trends of key phenotypic descriptors withing the PDI and DAI measures
- Agrilus exit hole density had no contribution
- Top ranked descriptor index response surfaces

Crown related descriptors were found to contribute most to the PDI (Figure #a) with more severely declining trees having reduced crown volume and condition. 
The DAI was a more complex in descriptor contribution with trunk staining and bleeding descriptors contributing highly with bleed prevalence contributing the most. 
Trees with high bleed prevalence and therefore a larger area of their trunk covered with black staining or active bleeds had much higher DAI scores.

The phenotypic indexes of all 171 trees (including the 20 trees Chestnuts) are shown in Figure 3. 
It can be seen that across the 9 sites, purely chronic oak decline sites such as Chestnuts and Speculation are easily distinguishable using the DAI compared to more AOD dominant sites such as Langdale or mixed sites such as Great Monks Wood.

The Phenotypic Decline Index (PDI) and Decline Acuteness Index (DAI) were previously defined in the 30 month report. 
A machine learning approach was used to derive these measures as it allows the complex interactions and trends between phenotypic descriptors to be modelled from which the decline indexes are calculated. 
This would likely be oversimplified using conventional modelling strategies. 
These complex interactions can be investigated by simulating phenotypic data and generating the index values from random forest models that have been trained for each of the indexes, supervised using phenotypic descriptors from the 171 trees. 
The index values generated by the models from these simulated phenotypic data can then be visualised as response surface plots, from which descriptor interactions within the indexes can be interpreted. 
Figures 1 and 2 show the simulated index value response surfaces under a variety of different circumstances for key phenotypic descriptors that contribute highly to the PDI and DAI.

@@

>>>>

### Descriptor contribution to decline indexes

@@

For the PDI, missing crown and crown transparency were compared for example cases of a healthy, a moderate and a severly tree.
The example cases included the trees with the minimum and maximum PDI values and the tree with the PDI value closest to the median for the healthy, severe and moderately declining trees respectively.
The value ranges for the missing crown (`r glue("{min(site_corrected_analysis_suitable_data[['Missing crown (%)']])}-{max(site_corrected_analysis_suitable_data[['Missing crown (%)']])}%")`) and crown transparency (`r glue("{min(site_corrected_analysis_suitable_data[['Crown transparency (%)']])}-{max(site_corrected_analysis_suitable_data[['Crown transparency (%)']])}%")`) descriptors divided into 100 increments.
For a given case, each combination of theses simulated descriptor values were used to calculate a PDI value from the supervised random forest regression model, building a model response surface.

@@

### Key descriptors contributing to the decline indexes

- Top ranked descriptors contributing to the decline indexes

The top ranked descriptors found to contr crown related descriptors were found to contribute most to the PDI (Figure #a) with more severely declining trees having reduced crown volume and condition.

- Plots of the individual descriptors available

The DAI was a more complex in descriptor contribution with trunk staining and bleeding descriptors contributing highly with bleed prevalence contributing the most. 
Trees with high bleed prevalence and therefore a larger area of their trunk covered with black staining or active bleeds had much higher DAI scores.

- Agrilus exit hole density had no contribution

### Key descriptor interactions within the decline index models

- Top ranked descriptor index response surfaces
- Trends of key phenotypic descriptors within the PDI and DAI measures
- Relatively smaller trees associated with more severe decline.

These complex interactions can be investigated by simulating phenotypic data and generating the index values from random forest models that have been trained for each of the indexes, supervised using phenotypic descriptors from the 171 trees. 
The index values generated by the models from these simulated phenotypic data can then be visualised as response surface plots, from which descriptor interactions within the indexes can be interpreted. 
Figures 1 and 2 show the simulated index value response surfaces under a variety of different circumstances for key phenotypic descriptors that contribute highly to the PDI and DAI.

@@

A machine learning approach was used to derive these measures as it allows the complex interactions and trends between phenotypic descriptors to be modelled from which the decline indexes are calculated. 
This would likely be oversimplified using conventional modelling strategies.
